On the reality of the quantum state once again: A no-go theorem for $\psi$-ontic models

### 2023-10-14 Articles/2021-HarriganSpekkens/HarriganSpekkensTakeTwo.tex
Changes from Andrea

\documentclass[10pt,twocolumn, nofootinbib]{revtex4-2}
%\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}
%\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{enumitem}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{tikz}

@@

In 2010 Harrigan and Spekkens (HS) proposed a rigorous classification in order to categorize the nature of the quantum state, i.e.\ to establish whether in a certain model $\psi$ corresponds to a real property of a quantum object, in which case the model is called $\psi$-ontic, or to some observer information, making it $\psi$-epistemic \cite{Harrigan:2010}. While their original aim was to clarify Einstein's view of Quantum Mechanics (QM), the HS framework has been widely employed in the literature not only to categorize different formulations of QM, but also to argue what types of interpretations are admissible (\cite{Leifer:2013, Leifer:2017, Branciard:2014, Hermens:2021, Wood:2015, Ringbauer:2015, Mazurek:2016, Bartlett:2012}; cf.\ \cite{Oldofredi:2020b, Ladyman:2021} for critical discussions).

@@

Such a theorem had a remarkable resonance (\cite{Leifer:2014, Leifer:2014b, Lewis:2012, Renner:2012, Colbeck:2017, Hardy:2013, Maroney:2014, Patra:2013, Mansfield:2016, Schlosshauer:2012, Schlosshauer:2013, Schlosshauer:2014, Aaronson:2013}), and questions about its actual meaning are still discussed today: on the one hand, it rules out interpretations of QM where $\psi$ merely represents information. On the other, scholars recently showed that non-trivial epistemic as well as statistical approaches to QM are not refuted by the PBR argument (\cite{Ben:2017, Rizzi:2018, Oldofredi:2021, DeBrota:2019}).

While the discussion has been focused mainly on whether $\psi$-epistemic models are problematic or whether the PBR argument itself is problematic, we ask a different question: what if the underlying HS classification itself is problematic? In particular, the framework assumes that once pure states are fully characterized, mixed states are constructed simply as a standard (i.e. classical) statistical mixture. However, as it is well-known, classical and quantum mixtures have different properties. For instance, classical mixtures have a single decomposition in terms of pure states while quantum mixtures do not. Using information theoretic considerations on mixtures, then, provides a new approach to explore possible violations of quantum mechanics.

Using this new approach, in this paper we show that $\psi$-ontic models cannot reproduce all the predictions, results and implications of QM. We start by noting that the von Neumann entropy plays a crucial role in the predictions of both quantum statistical mechanics and quantum information theory. We show that the degree of overlap between two probability distributions affects how the entropy of their mixture relates to the entropy of the components. We find that the lack of overlap of epistemic states---a necessary condition for a model to be $\psi$-ontic---requires that all pure states must be orthogonal, which directly contradicts quantum theory.

@@

To classify an ontological model, we look at the relationship between quantum and ontic states. There are two broad categories. In a \textbf{$\psi$-ontic} model the wave function is a physical property of the ontic state, in the sense that given an ontic state $\lambda$ there is only one pure state preparation $P_\psi$ that could have prepared $\lambda$. As we can see in figure \ref{overlap} (a), this happens if the probability distributions do not overlap, i.e.\ if we have
\begin{equation}\label{ontic_condition}
	p(\lambda | P_{\psi})p(\lambda|P_{\phi})=0
\end{equation}
for all pairs of states $\psi$ and $\phi$. If a model is not $\psi$-ontic, then it is \textbf{$\psi$-epistemic}. In this case, $\lambda$ can be described by more than one quantum state and the wave function is taken to represent knowledge about the state preparation---in such models quantum states generate overlapping probability distributions over $\Lambda$ as shown in figure \ref{overlap} (b). A $\psi$-ontic model is said \textbf{$\psi$-complete} if the quantum states and the ontic states coincide. More precisely if 
\begin{equation}\label{complete_condition}
	p(\lambda|P_\psi)=\delta(\lambda-\lambda_{\psi}).
\end{equation}
All other models are \textbf{$\psi$-incomplete}. For specific examples of ontological models, see \cite{Harrigan:2010}, sections 2.4.1--3.

@@

@@

Let us denote $M(\Lambda)$ the space of all probability distributions (i.e. all probability measures) over the space of ontic states $\Lambda$. Let us denote $\mathcal{H}$ the Hilbert space of the corresponding quantum system and $M(\mathcal{H})$ the space of all mixed states (i.e. the space of positive semi-definite trace-one Hermitian operators). An ontological model, then, must give us an appropriate $\rho(\lambda) \in M(\Lambda)$ for every $\rho \in M(\mathcal{H})$ such that all predictions are satisfied. This includes those given by statistical mechanics and information theory, which depend on the von Neumann entropy $H_\mathcal{H} : M(\mathcal{H}) \to \mathbb{R}$. Given that $M(\Lambda)$ is the space of probability measures over $\Lambda$, it is natural to use the Shannon/Gibbs entropy function $H_\Lambda : M(\Lambda) \to \mathbb{R}$, as this is what is typically done in statistical mechanics.\footnote{Note that, while many functions called ``entropy'' exist, the Shannon entropy is the only indicator of variability that is continuous, monotonic and linear in probability.\cite{aop-phys-variability} It is also the one that provides the correct link to the thermodynamic entropy.} Since we want to be able to replicate the predictions of statistical mechanics and information theory, a very reasonable request is that these two entropies, calculated on the different representations, agree.

We now introduce two lemmas that show how the entropy behaves when mixing two states, both for the Shannon/Gibbs entropy $H_\Lambda$ and the von Neumann entropy $H_\mathcal{H}$.\footnote{Full calculations with all steps are included in the appendix.} The first will be used to calculate entropy of statistical mixtures of epistemic states, since equation \ref{epistemic_mixing} tells us the mixture is done according to a standard (classical) measure of probability. The second will be used to calculate the entropy on the quantum case.

@@

\textbf{Quantum information entropy of quantum mixed states}. Now suppose $\psi$ and $\phi$ are two pure quantum states and let $p = | \langle \psi | \phi \rangle |^2$ be the probability of transition from one to the other. Consider the mixed state $\rho = \frac{1}{2} | \psi \rangle \langle \psi | + \frac{1}{2} | \phi \rangle \langle \phi |$. Its entropy is given by
\begin{equation}\label{entropy_mixed}
	\begin{aligned}
	H_\mathcal{H}(\rho) &= H_S\left(\frac{1+\sqrt{p}}{2}, \frac{1-\sqrt{p}}{2}\right) \\
	&= - \frac{1+\sqrt{p}}{2} \log \frac{1+\sqrt{p}}{2} - \frac{1-\sqrt{p}}{2} \log \frac{1-\sqrt{p}}{2} ,
	\end{aligned}
\end{equation}
where the right side of the equation is the Shannon entropy calculated on the given values.\footnote{Note that $H_S$ is the ``discrete'' Shannon entropy that acts on a countable set of coefficients that sum to one.} Therefore the entropy of an equal mixture of two pure states depends only on the probability of measuring one having prepared the other.

\textbf{Theorem: Since non-overlapping distributions can only represent orthogonal states, $\psi$-ontic models cannot be consistent with quantum theory.}\footnote{Note that the theorem applies to all $\psi$-ontic models, including the $\psi$-supplemented ones.}

@@

First of all, information entropy, in all its versions, is strictly concave, meaning that $H(p \rho_1 + (1-p) \rho_2) \geq p H(\rho_1) + (1-p) H(\rho_2)$ with the equality valid only if $\rho_1$ and $\rho_2$ are the same exact state. That is, we can check whether two states $\rho_1$ and $\rho_2$ are the same state simply by creating a mixture $\rho = \frac{1}{2} \rho_1 + \frac{1}{2} \rho_2$ and see whether we end up with the same entropy or not. In other words: \emph{entropy fixes equality}.
The issue here is that the mapping between epistemic states and density matrices is not injective. Quantum mixtures, in fact, do not have a single decomposition in pure states. For example, the maximally mixed state of a qubit for a spin-$1/2$ system can be achieved with either an equal mixture of spin up and down or an equal mixture of spin left and right. Therefore epistemic states cannot be mapped one-to-one to quantum density operators, but rather each density operator corresponds to an equivalence class of epistemic states.\footnote{The degree of degeneracy one has to require is also extreme. For a qubit, in the simplest case, we would have a $\psi$-ontic state for each direction on the Bloch sphere, meaning the space of ontic states is the sphere $S^2$. The space of all possible epistemic states is the set of all possible distributions integrable to one. This is an infinite dimensional function space. However, the space of mixed states for a quibit is the interior of the Bloch sphere, which is a bounded three-dimensional manifold. The equivalence classes of epistemic states, then, must be infinite dimensional.} This degeneracy means that the notion of equality among epistemic states is different than the notion of equality among density operators. In turn, this means that whatever definition we have for entropy on epistemic states must not satisfy strict concavity, if we want the value of entropy to agree between epistemic states and density operators. \emph{Entropy cannot fix equality for epistemic states}. In other words, one can make mixtures of different epistemic states without increasing entropy, which is questionable on physics grounds.

@@

In our result, we saw that the use of standard measure theory when creating mixtures as in equation \ref{epistemic_mixing} generates a problem because the convex space of classical distributions is significantly different from the convex space of quantum mixed states. For this case, ontological models are ``not quantum enough.''

In PBR, the issue is breaking a composite system into parts. They find  that if the probability measure on $\lambda$ is factorized, that is (\cite{PBR:2012}, eq. 5) 
\begin{equation}
	\int_{\Lambda} \dots \int_{\Lambda} p(k | \lambda_1, \cdots, \lambda_n) \mu_{x_1}(\lambda_1)\cdots\mu_{x_n}(\lambda_n) d\lambda_1 \cdots d\lambda_n ,
\end{equation}
there is a problem. However, the issue here is not the factorization, but rather breaking the composite system into parts, each with its own ontic state as in 
\begin{equation}
	\begin{aligned}
	\int_\Lambda &p(k | \lambda) d\lambda = \\
	&\int_{\Lambda_1} \dots \int_{\Lambda_n} p(k | \lambda_1, \cdots, \lambda_n) \mu(\lambda_1, \cdots, \lambda_n) d\lambda_1 \cdots d\lambda_n.
	\end{aligned}
\end{equation}
This assumes that a composite system is describable using a standard classical probability measure, which, in turn, allows us to express the joint probability distribution of observables of different subsystems. This, in general, will fail because it is ``not quantum enough.''

@@

Now, the fact that quantum theory does not follow the rules of classical-Kolmogorov probability is nothing new\cite{Werner:2014}. However, this does not seem to be enough to clear confusion on the subject because, admittedly, we do use classical probability in the context of quantum mechanics. For example, we can imagine preparing the direction of spin based on a classical distribution. Furthermore, the output of a measurement is described by classical probability. The temptation, then, is to simply assume that all we need to do is just put the right transition probability between the two spaces, and we are done. This is exactly what cannot be done and what doesn't work. That is
\begin{quote}
	\textbf{In the context of quantum mechanics, standard probability measures are allowed over preparations and measurement outcomes, but not over states.}
\end{quote}
In our approach, we exploited the fact that the entropy calculation uses the geometry of the inner product without invoking a probability of transition. That is, two pure states are orthogonal not because one cannot measure the first having prepared the second, but because their equal mixture raises the entropy by one bit. Therefore quantum mechanics is not simply defining a probability of transition between preparations and measurements: it is doing something more. TODO: We hope that this additional insight may provide more precise guidance as to where classical probability calculus is appropriate or not in the context of quantum mechanics.

@@

The key problem of the HS categorization is that it is ``not quantum enough.'' The framework tries to hide the non-classicality at the level of each ontic state, which, by construction, cannot be decomposed further using a standard probability measure. However, HS definitions entail that statistical mixtures are simply classical mixtures in the sense that they are captured by a standard probability measure. But quantum mixtures are non-classical in the same way that the quantum transition probabilities defined by the inner product are non-classical.\footnote{The convex space formed by the mixtures is radically different\cite{bengtsson_życzkowski_2017}. There is also debate as whether proper and improper mixtures are distinguishable in the context of quantum mechanics\cite{Kirkpatrick:2001,DEspagnat:2001}.} Therefore, the HS categorization prevents one from describing quantum statistical mixing.

Our conclusion is that the PBR theorem finds that $\psi$-epistemic models are untenable not because there is a problem with $\psi$-epistemic models in particular, but rather there is an underlying problem in the HS definitions. Specifically, the problem of the HS categorization is the use of standard probability theory in composite systems. If understood in this manner, the PBR result is the first part of a more general argument showing the inadequacy of the HS framework to categorize quantum theories, an argument that we have completed with the proposed no-go theorem against $\psi$-ontic models.

As a final note, one may argue that our result cannot possibly be correct because there are actual examples of $\psi$-ontic models. To our knowledge, however, none of them is able to reproduce all results from quantum statistical mechanics (e.g. derive the Boltzmann distribution from entropy maximization) or quantum information theory (e.g. derive the Holevo bound). Since the latter are derived from the more general framework of quantum mechanics, one cannot cherry pick which results a model has to reproduce to be a valid model of quantum theory: \emph{all} results must be correctly reproduced.

>>>>

\documentclass[10pt,twocolumn, nofootinbib]{revtex4-2}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{enumitem}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{tikz}

@@

In 2010 Harrigan and Spekkens (HS) proposed a \textcolor{magenta}{formal} classification in order to categorize the nature of the quantum state, i.e.\ to establish whether in a certain model $\psi$ corresponds to a real property of a quantum object, in which case the model is called $\psi$-ontic, or to some observer information, making it $\psi$-epistemic \cite{Harrigan:2010}. While their original aim was to clarify Einstein's view of Quantum Mechanics (QM), the HS framework has been widely employed in the literature not only to categorize different formulations of QM, but also to argue what types of interpretations are admissible (\cite{Leifer:2013, Leifer:2017, Branciard:2014, Hermens:2021, Wood:2015, Ringbauer:2015, Mazurek:2016, Bartlett:2012}; cf.\ \cite{Oldofredi:2020b, Ladyman:2021} for critical discussions).

@@

Such a theorem had a remarkable resonance (\cite{Leifer:2014, Leifer:2014b, Lewis:2012, Renner:2012, Colbeck:2017, Hardy:2013, Maroney:2014, Patra:2013, Mansfield:2016, Schlosshauer:2012, Schlosshauer:2013, Schlosshauer:2014, Aaronson:2013}), and questions about its actual meaning are still discussed today: on the one hand, \textcolor{magenta}{some authors believe that it rules out interpretations of QM where $\psi$ merely represents information. On the other hand, it has been recently showed by other scholars that non-trivial epistemic as well as statistical approaches to QM are not refuted by the PBR argument (\cite{Ben:2017, Rizzi:2018, Oldofredi:2020b, Oldofredi:2021, DeBrota:2019}).}

While the discussion has been focused mainly on whether $\psi$-epistemic models are problematic or whether the PBR argument itself is problematic, we ask a different question: what if the underlying HS classification itself is problematic? In particular, the framework assumes that once pure states are fully characterized, mixed states are constructed simply as a standard (i.e.\ classical) statistical mixture. However, as it is well-known, classical and quantum mixtures have different properties. For instance, classical mixtures have a single decomposition in terms of pure states while quantum mixtures do not. Using information theoretic considerations on mixtures, then, provides a new approach to explore possible violations of quantum mechanics.

\textcolor{magenta}{Employing} this new approach, in this paper we show that $\psi$-ontic models cannot reproduce all the predictions, results and implications of QM. \textcolor{blue}{We start by noting that the von Neumann entropy plays a crucial role in the predictions of both quantum statistical mechanics and quantum information theory. We show that} the degree of overlap between two probability distributions affects how the entropy of their mixture relates to the entropy of the components. We find that the lack of overlap of epistemic states---a necessary condition for a model to be $\psi$-ontic---requires that all pure states must be orthogonal, which directly contradicts quantum theory.

@@

To classify an ontological model, we look at the relationship between quantum and ontic states. There are two broad categories. In a \textbf{$\psi$-ontic} model the wave function is a physical property of the ontic state, in the sense that given an ontic state $\lambda$ there is only one pure state preparation $P_\psi$ that could have prepared $\lambda$. As we can see in figure \ref{overlap} (a), this happens if the probability distributions do not overlap, i.e.\ if we have
\begin{equation}\label{ontic_condition}
	p(\lambda | P_{\psi})p(\lambda|P_{\phi})=0
\end{equation}
for all pairs of states $\psi$ and $\phi$. If a model is not $\psi$-ontic, then it is \textbf{$\psi$-epistemic}. In this case, $\lambda$ can be described by more than one quantum state and the wave function is taken to represent knowledge about the state preparation---in such models quantum states generate overlapping probability distributions over $\Lambda$ as shown in figure \ref{overlap} (b). A $\psi$-ontic model is said \textbf{$\psi$-complete} if the quantum states and the ontic states coincide. More precisely if 
\begin{equation}\label{complete_condition}
	p(\lambda|P_\psi)=\delta(\lambda-\lambda_{\psi}).
\end{equation}
All other models are \textbf{$\psi$-incomplete}. \textcolor{blue}{For specific examples of ontological models, see \cite{Harrigan:2010}, sections 2.4.1--3.}

@@

\textcolor{magenta}{After having introduced the relevant definitions of the HS ontological model framework, it should be noted that it has been subject to several criticisms, as already said in the previous section. For instance, Oldofredi \& Lopez in \cite{Oldofredi:2020b} analyzed the implicit assumptions made by Harrigan \& Spekkens in order to define their categorization, and showed that relevant interpretations of quantum theory---such as the statistical, relational and perspectival readings of the quantum formalism---cannot be classified according to the HS approach, given that their principles are in tension with those employed in \cite{Harrigan:2010}. Similarly, Hance, Rarity and Ladyman argued in \cite{Ladyman:2021} that the dichotomy ``ontic vs.\ epistemic'' proposed by HS is artificial, since there are cases in which the quantum state can represent both ontic properties of quantum systems as well as agents' knowledge of them. Thus, they claim, the definitions of $\psi$-ontic and $\psi$-epistemic introduced a few lines above do not capture the intuitive ideas with which physicists and philosophers alike use the terms ``ontic'' and ``epistemic''.}

\textcolor{magenta}{While we agree with the main messages of these critical responses to the HS framework, the argument offered in this essay follows a different and complementary strategy. In fact, instead of criticizing the assumptions used by Harrigan \& Spekkens, we consider them valid (for the sake of the argument) and derive a no-go theorem for $\psi$-ontic models directly from them, as we are going to show in the next section.}

@@

\textcolor{blue}{Let us denote $M(\Lambda)$ the space of all probability distributions (i.e. all probability measures) over the space of ontic states $\Lambda$. Let us denote $\mathcal{H}$ the Hilbert space of the corresponding quantum system and $M(\mathcal{H})$ the space of all mixed states (i.e. the space of positive semi-definite trace-one Hermitian operators). An ontological model, then, must give us an appropriate $\rho(\lambda) \in M(\Lambda)$ for every $\rho \in M(\mathcal{H})$ such that all predictions are satisfied. This includes those given by statistical mechanics and information theory, which depend on the von Neumann entropy $H_\mathcal{H} : M(\mathcal{H}) \to \mathbb{R}$. Given that $M(\Lambda)$ is the space of probability measures over $\Lambda$, it is natural to use the Shannon/Gibbs entropy function $H_\Lambda : M(\Lambda) \to \mathbb{R}$, as this is what is typically done in statistical mechanics.\footnote{\textcolor{blue}{Note that, while many functions called ``entropy'' exist, the Shannon entropy is the only indicator of variability that is continuous, monotonic and linear in probability \cite{Carcassi:2021}. It is also the one that provides the correct link to the thermodynamic entropy.}} Since we want to be able to replicate the predictions of statistical mechanics and information theory, a very reasonable request is that these two entropies, calculated on the different representations, agree.}

\textcolor{blue}{We now introduce two lemmas that} show how the entropy behaves when mixing two states, both for the Shannon/Gibbs entropy $H_\Lambda$ and the von Neumann entropy $H_\mathcal{H}$.\footnote{Full calculations with all steps are included in the appendix.} The first will be used to calculate entropy of statistical mixtures of epistemic states, since equation \ref{epistemic_mixing} tells us the mixture is done according to a standard (classical) measure of probability. The second will be used to calculate the entropy on the quantum case.

@@

\textbf{Quantum information entropy of quantum mixed states}. Now suppose $\psi$ and $\phi$ are two pure quantum states and let $p = | \langle \psi | \phi \rangle |^2$ be the probability of transition from one to the other. Consider the mixed state $\rho = \frac{1}{2} | \psi \rangle \langle \psi | + \frac{1}{2} | \phi \rangle \langle \phi |$. Its entropy is given by
\begin{equation}\label{entropy_mixed}
	\begin{aligned}
	H_\mathcal{H}(\rho) &= H_S\left(\frac{1+\sqrt{p}}{2}, \frac{1-\sqrt{p}}{2}\right) \\
	&= - \frac{1+\sqrt{p}}{2} \log \frac{1+\sqrt{p}}{2} - \frac{1-\sqrt{p}}{2} \log \frac{1-\sqrt{p}}{2} ,
	\end{aligned}
\end{equation}
where the right side of the equation is the Shannon entropy calculated on the given values.\footnote{\textcolor{blue}{Note that $H_S$ is the ``discrete'' Shannon entropy that acts on a countable set of coefficients that sum to one.}} Therefore the entropy of an equal mixture of two pure states depends only on the probability of measuring one having prepared the other.

\textbf{Theorem: Since non-overlapping distributions can only represent orthogonal states, $\psi$-ontic models cannot be consistent with quantum theory.}\footnote{\textcolor{blue}{Note that the theorem applies to all $\psi$-ontic models, including the $\psi$-supplemented ones.}}

@@

First of all, information entropy, \textcolor{blue}{in all its versions,} is strictly concave, meaning that $H(p \rho_1 + (1-p) \rho_2) \geq p H(\rho_1) + (1-p) H(\rho_2)$ with the equality valid only if $\rho_1$ and $\rho_2$ are the same exact state. That is, we can check whether two states $\rho_1$ and $\rho_2$ are the same state simply by creating a mixture $\rho = \frac{1}{2} \rho_1 + \frac{1}{2} \rho_2$ and see whether we end up with the same entropy or not. In other words: \emph{entropy fixes equality}.
The issue here is that the mapping between epistemic states and density matrices is not injective. Quantum mixtures, in fact, do not have a single decomposition in pure states. For example, the maximally mixed state of a qubit for a spin-$1/2$ system can be achieved with either an equal mixture of spin up and down or an equal mixture of spin left and right. Therefore epistemic states cannot be mapped one-to-one to quantum density operators, but rather each density operator corresponds to an equivalence class of epistemic states.\footnote{The degree of degeneracy one has to require is also extreme. For a qubit, in the simplest case, we would have a $\psi$-ontic state for each direction on the Bloch sphere, meaning the space of ontic states is the sphere $S^2$. The space of all possible epistemic states is the set of all possible distributions integrable to one. This is an infinite dimensional function space. However, the space of mixed states for a quibit is the interior of the Bloch sphere, which is a bounded three-dimensional manifold. The equivalence classes of epistemic states, then, must be infinite dimensional.} This degeneracy means that the notion of equality among epistemic states is different than the notion of equality among density operators. In turn, this means that whatever definition we have for entropy on epistemic states must not satisfy strict concavity, if we want the value of entropy to agree between epistemic states and density operators. \emph{Entropy cannot fix equality for epistemic states}. In other words, one can make mixtures of different epistemic states without increasing entropy, which is questionable on physics grounds.

@@

In our result, we saw that the use of standard measure theory when creating mixtures as in equation \ref{epistemic_mixing} generates a problem because the convex space of classical distributions is significantly different from the convex space of quantum mixed states. For this case, ontological models are ``not quantum enough''.

In PBR, the issue is breaking a composite system into parts. They find  that if the probability measure on $\lambda$ is factorized, that is (\cite{PBR:2012}, eq. 5) 
\begin{equation}
	\int_{\Lambda} \dots \int_{\Lambda} p(k | \lambda_1, \cdots, \lambda_n) \mu_{x_1}(\lambda_1)\cdots\mu_{x_n}(\lambda_n) d\lambda_1 \cdots d\lambda_n ,
\end{equation}
there is a problem. However, the issue here is not the factorization, but rather breaking the composite system into parts, each with its own ontic state as in 
\begin{equation}
	\begin{aligned}
	\int_\Lambda &p(k | \lambda) d\lambda = \\
	&\int_{\Lambda_1} \dots \int_{\Lambda_n} p(k | \lambda_1, \cdots, \lambda_n) \mu(\lambda_1, \cdots, \lambda_n) d\lambda_1 \cdots d\lambda_n.
	\end{aligned}
\end{equation}
This assumes that a composite system is describable using a standard classical probability measure, which, in turn, allows us to express the joint probability distribution of observables of different subsystems. \textcolor{magenta}{Again, this in general will fail because it is ``not quantum enough''.}

@@

\textcolor{blue}{Now, the fact that quantum theory does not follow the rules of classical-Kolmogorov probability is nothing new \cite{Werner:2014, Pitowsky:1989}. However, this does not seem to be enough to clear confusion on the subject because, admittedly, we do use classical probability in the context of quantum mechanics. For example, we can imagine preparing the direction of spin based on a classical distribution. Furthermore, the output of a measurement is described by classical probability. The temptation, then, is to simply assume that all we need to do is just put the right transition probability between the two spaces, and we are done. This is exactly what cannot be done and what does not work. That is}
\begin{quote}
	\textcolor{blue}{\textbf{In the context of quantum mechanics, standard probability measures are allowed over preparations and measurement outcomes, but not over states.}}
\end{quote}
\textcolor{blue}{In our approach, we exploited the fact that the entropy calculation uses the geometry of the inner product without invoking a probability of transition. That is, two pure states are orthogonal not because one cannot measure the first having prepared the second, but because their equal mixture raises the entropy by one bit. Therefore quantum mechanics is not simply defining a probability of transition between preparations and measurements: it is doing something more. While it is not in the scope of this article to articulate precisely what this ``more'' is, which will be the focus of future work, we hope that this additional insight may provide more precise guidance as to where classical probability calculus is appropriate and not in the context of quantum mechanics.}

@@

The key problem of the HS categorization is that it is ``not quantum enough''. The framework in fact tries to hide the non-classicality at the level of each ontic state which, by construction, cannot be decomposed further using a standard probability measure. However, HS definitions entail that statistical mixtures are simply classical mixtures in the sense that they are captured by a standard probability measure. But quantum mixtures are non-classical in the same way that the quantum transition probabilities defined by the inner product are non-classical.\footnote{\textcolor{blue}{The convex space formed by the mixtures is radically different \cite{Bengtsson:2017}. There is also debate as whether proper and improper mixtures are distinguishable in the context of quantum mechanics \cite{Kirkpatrick:2001}, \cite{DEspagnat:2001}.}} Therefore, the HS categorization prevents one from describing quantum statistical mixing.

Our conclusion is that the PBR theorem finds that $\psi$-epistemic models are untenable not because there is a problem with epistemic models in particular, but rather there is an underlying problem in the HS definitions. Specifically, the problem of the HS categorization is the use of standard probability theory in composite systems. If understood in this manner, the PBR result is the first part of a more general argument showing the inadequacy of the HS framework to categorize quantum theories, an argument that we have completed with the proposed no-go theorem against $\psi$-ontic models.

\textcolor{magenta}{Referring to this, two final remarks are in order: on the one hand, our analysis entails an interesting consequence for the philosophical discussion concerning the interpretation of quantum theory. Indeed, many scholars use the PBR result to vindicate the empirical adequacy of a given framework---especially those in which the quantum state is somehow supposed to represent real physical objects---or to rule out a specific reading of the quantum formalism. Here we warn physicists and philosophers that no argument resting on Harrigan \& Spekkens approach should be employed in such metaphysical debates, given that no theory satisfying the requirements imposed by their categorization is able to reproduce \emph{all} the results of quantum mechanics.}

\textcolor{magenta}{To this regard, however, one may argue that our result cannot possibly be correct because there are actual examples of $\psi$-ontic models. To our knowledge, however, none of them is fully empirically equivalent to quantum theory, i.e.\ is able to reproduce all results from quantum statistical mechanics (e.g., derive the Boltzmann distribution from entropy maximization) or quantum information theory (e.g., derive the Holevo bound). Since the latter are derived from the more general framework of quantum mechanics, one cannot cherry pick which results a model has to reproduce to be a valid model of quantum theory: \emph{all} results must be correctly reproduced. In sum, we not only agree with Hance, Rarity and Ladyman in saying that HS definitions of $\psi$-ontic and $\psi$-epistemic models do not capture the actual meaning usually assigned to the terms ``ontic'' and ``epistemic'' as they are employed when interpreting quantum theories, but also say the following:}
\begin{quote}
\textcolor{magenta}{\textbf{If a certain interpretation of the quantum formalism\footnote{\textcolor{magenta}{As for instance de Broglie-Bohm theory, Everett's relative state formulation, Relational quantum mechanics or the Many-Worlds perspective \emph{etc.}.}} is fully empirically equivalent to quantum mechanics, then it cannot be categorized as a $\psi$-ontic model.}}
\end{quote}
\noindent \textcolor{magenta}{In turn, the logical conjunction of our result and the PBR theorem entails that no interpretation of quantum theory can be correctly categorized according to the HS approach.}

\textcolor{magenta}{On the other hand, and we conclude, another philosophical issue affecting the definition of $\psi$-ontic model must be mentioned. In their essay Harrigan \& Spekkens require that if a model is $\psi$-ontic, the ontic state $\lambda$ is put into a  one-to-one correspondence with the quantum state, i.e.\ only one quantum state is compatible with the ontic state of the system. This fact, in turn, means that $\psi$ should be understood realistically, as an entity representing some real physical property of the system, or even the system itself as in $\psi$-complete models. However, as interestingly noted by Halvorson, realism towards the quantum state is highly problematic, since each representation relation proposed in literature does entail several conundrums that should be avoided---cf.\  \cite{Halvorson:2019} for details. To this regard, in fact, HS do not clearly spelled out what they mean when it is asserted that $\psi$ ``represents'' a real property of physical systems. Alternatively, to claim that $\psi$-ontic models $\lambda$ is uniquely represented by a single quantum state leaves many ontological questions unanswered: for instance it is not specified in which way $\psi$ represents the ontic state, is it a direct or indirect representation? If $\psi$ is a real object, as one would expect in $\psi$-complete models, is there a literal identification between $\psi$ and $\lambda$? To these and other issues Harrigan \& Spekkens do not reply.}

\textcolor{magenta}{Therefore, not only their categorization is shown to be empty, but also it does not offer a rigorous philosophical ground where to cast the interpretational debate.}
